\LARGE How coexistence mechanisms mediate temporal stability

### 2014-10-17 modelNotes.Rnw
working storage effect mode in cont time!
this, v2, is the working version. Can delete original.
updated equations to match *v2 storage model

\subsection{Temporal storage effect}

>>>>

\subsection{Temporal storage effect}
We start with the continuous-time model

\begin{align}
\frac{\text{d}D_{i}}{\text{d}t} &= c_{i}b_{i}(N_{i} + g_{i,e}D_{i}) - m_{D,i}D_{i}\\
\frac{\text{d}N_{i}}{\text{d}t} &= f(R)(N_{i} + g_{i,e}D_{i}) - c_{i}b_{i}(N_{i} + g_{i,e}D_{i})  - m_{N,i}(N_{i}+g_{i,e}D_{i})\\
\frac{\text{d}R_{i}}{\text{d}t} &= a(S - R) - \sum\limits_{i=1,2}f(R)(N_{i}+g_{i,e}D_{i})
\end{align}
where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, $B$ is the living biomass (fast-growing, shorter-lived) state, $b$ is the ``birth" rate of new biomass in the dormant state, $c$ is some factor related to the cost of producing the dormant state, and $m$s are biomass loss rates. $g$ is a time-fluctuating activation rate of dormant biomass to the live biomass state that occurs at discrete times $e$ (in our case, every `year'). The growth rate of living biomass is a resource-dependent function, $f(R) = r_{i}R/(K_{i}+R)$, where $r$ is the maximum growth rate and $K$ is the resource level at which growth is one-half $r$. For the resource dynamics, whose state is denoted by $R$, we use a linear resource renewal equation where $a$ scales resource turnover rate and $S$ is the resource equilibrium when consumers are absent, and an offtake of the resource equal to the sum of each species' consumption, $\sum_{i=1,2}f(R)(N_{i}+g_{i,e}D_{i})$.

To make this a ``storage-effect" model, we need to satisfy three conditions: (1) the organisms must have a mechanism for persistence under unfavorable conditions, (2) species must respond differently to environmental conditions, and (3) the effects of competition on a species must be more strongly negative in good years relative to unfavorable years. Our model meets condition 1 because we include a dormant stage with very low death rates. We satisfy condition 2 with our model whenever $g$ is not perfectly correlated between species. Lastly, our model meets condition 3 because condition 2 partitions intraspecific and interspecific competition into different years. Thus, during a high $g$ year for one species, resource uptake is also inherently high for that species, which increases intraspecific competition relative to interspecific competition. So, given adequate variability in $g$, the inferior competitor (species with lower $r$) can persist.
